An Investigation of Confusing Code Patterns in JavaScript

### 2022-12-02 s03.tex
changing the paper organization

%% \section{Results and Discussion}
%% \label{sec:results}

%% In this section we present the main findings of our research. We first present the results of each study (Section~\ref{sec:survey-resuts}, Sections~\ref{sec:msr-results}, and Section~\ref{sec:interview-results}). After that, in Section~\ref{sec:discussion} we consolidate our findings and present some implications of our study.

%\section{Results of the Survey}
%\label{sec:survey-resuts}

%-----------------------------------------------------------------------
\section{Results of the Repeated Measures Study}\label{sec:repeated}

In this experiment, we investigate whether 24 atom candidates impact readability. We examine how the performance of the participants differ when predicting the outputs of the clean and obfuscated versions of small code snippets (Section~\ref{sec:repeated:correct}). Based on those results we establish which candidates can actually be considered atoms of confusion. In addition, similarly to Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, we verify whether there is a trade-off between correctness and time irrespective of the presence of atoms; we check if participants who correctly predict the outputs of the programs tend to do so more slowly than the ones that make incorrect predictions (Section~\ref{sec:repeated:time}).

This study had 70 participants. Out of these, 67 have received university-level training in programming, 4 hold a PhD degree, 7 hold a master's degree, and 11 hold a bachelor's degree. 
The median programming experience of the participants is four years (mean 6.5 years) and the median experience with JavaScript programming is one year (mean 2.5 years). 
% General programming: median: 48, 1st quartile: 22,5, 3rd quartile: 91,5
% JavaScript programming: median: 12, 1st quartile: 4, 3rd: 24

%-----------------------------------------------------------------------
\subsection{Correctness Analysis}\label{sec:repeated:correct}

\subsubsection*{Exploratory Data Analysis}
For each of the 24 atom candidates, the participants predicted the outputs of two versions, one clean (control) and one obfuscated (treatment), as per the definition of a repeated measures design~\cite{Hunter-Experimenters}. Thus, each participant examined 48 code snippets.  We also measured the time required by the participants to provide their answers. For most cases, participants made more mistakes when attempting to predict the outcomes of the obfuscated versions.

Table~\ref{tab:diff-correctness-repeated} presents the number of correct predictions for the clean and obfuscated versions of each atom candidate, as well as the percentage difference. In only three cases the participants predicted results correctly more often for the obfuscated versions, Arithmetic as Logic, Logic as Control Flow, and Property Access. The former two were empirically validated as atoms of confusion in the study of Gopstein and colleagues~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For seven of the atom candidates, the number of correct predictions for the clean versions is more than twice the corresponding number for the obfuscated versions.  Some of these atoms are common programming patterns in JavaScript programs, such as pre- and post-increments and object destructuring. The difference was even greater for Type Conversion (+440\%) and Change of Literal Encoding (+300\%).

\begin{table}[t!]
\caption{Summary of the correctness analysis.}
\label{tab:diff-correctness-repeated}
\centering{
  \begin{scriptsize}
\begin{tabular}{lrrr} \toprule
  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
Type Conversion           &            5 &      27 &                 +440 \\
Change of Literal Encoding &           9 &      36 &                 +300 \\
Comma Operator            &            8 &      29 &                 +262 \\
Object Destructuring      &           15 &      42 &                 +180 \\
Assignment As Value       &           18 &      43 &                 +139 \\
Pre-Increment             &           13 &      28 &                 +115 \\
Post-Increment            &            8 &      17 &                 +112 \\
Array Destructuring       &           15 &      28 &                 +87 \\
Lack of Indentation no Braces     &           30 &      50 &                 +67 \\
Omitted Curly Braces      &           35 &      51 &                 +46 \\
Object Spread             &           21 &      30 &                 +43 \\
Array Spread              &           22 &      31 &                 +41 \\
Automatic Semicolon Insertion &       15 &      21 &                 +40 \\
Infix Operator Precedence &           47 &      53 &                 +13 \\
Arrow Function            &           54 &      58 &                  +7 \\
Implicit Predicate        &           41 &      44 &                  +7 \\
Lack of Indentation With Braces   &           53 &      55 &                  +4 \\
Conditional Operator      &           60 &      62 &                  +3 \\
Constant Variables        &           59 &      61 &                  +3 \\
Dead Unreachable Repeated &           63 &      64 &                  +2 \\
Repurposed Variables      &           23 &      23 &                   0 \\
Arithmetic as Logic       &           52 &      47 &                 -10 \\
Logic as Control Flow     &           33 &      28 &                 -15 \\
Property Access           &           51 &      42 &                 -18 \\
\bottomrule  
\end{tabular}
\end{scriptsize}
}
\end{table}

\subsubsection*{Statistical analysis}

Since this study uses a within-subject design, the data is paired and dependent. Our goal is to verify whether there is a significant difference in the performance of the same participants when examining clean and obfuscated versions of programs. The analyzed data is also dichotomous, since each participant either correctly predicts the output of a program or does not. Considering this scenario, we employ McNemar's test~\cite{McNemar:1947:NSE}, which is aimed at \textit{``judging the significance of the difference between correlated proportions''}.

The ``McNemar test'' column of Table~\ref{tab:repeated:misunderstanding} indicates the p-values for the tests. These p-values are compared against an alpha of 0.0021, after applying Bonferroni correction (0.05/24). One of the atom candidates, Constant Variables, is omitted from the table because, due to the very low number of mistakes, it is not possible to perform the test. The table shows that eight of the 24 analyzed atom candidates can be considered atoms of confusion based on the results of our study, Object Destructuring, Type Conversion, Change of Literal Encoding, Comma Operator, Lack of Indentation No Braces, Assignment as Value, Pre-Increment, and Ommitted Curly Braces. Among these eight, only Object Destructuring is JavaScript-specific. Six of the other atoms have been previously reported for the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} and six for Java~\cite{Langhout:2021:ACJ}.
% Lack of Indentation No Braces is not specific to JavaScript and has also been studied by Langhout and Aniche~\cite{Langhout:2021:ACJ}. 
For some candidates that have been previously identified as atoms, there was not a statistically significant difference. This is the case for Repurposed Variables, Logic as Control Flow, Implicit Predicate, and Conditional Operator. We examine this in Section~\ref{sec:discussion}.

We leverage the odds ratio (OR) as a measure of effect size, to quantify the meaningfulness~\cite{Ellis:2010:EGE} of these results. The odds ratio is an appropriate measure of effect size for scenarios where McNemar's test is applicable~\cite{Mangiafico:2016:SAE}. For example, the odds ratio for Type Conversion is 12. This suggests that the odds of a participant correctly predicting the output of a clean version while incorrectly predicting the output of the corresponding obfuscated version is 12 times higher than the odds of correctly predicting the output of the obfuscated version and incorrectly predicting the output of the correct one.

According to the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for most of the atoms where there is a statistically significant difference the odds ratio can be considered high. The value of the odds ratio for Object Destructuring is $\infty$ because no participant correctly guessed the output for the obfuscated version while incorrectly guessing the output for the clean one, which results in a denominator of 0. Omitted Curly Braces exhibits an OR of 6.3, which is borderline between medium and high~\cite{Chen:2010:HBB}. Array Destructuring and Post-Increment are two atom candidates that deserve a special note. The difference in the performance of participants examining clean and obfuscated versions of these atoms was not statistically significant, after correction. Notwithstanding, effect sizes were large and medium, respectively. This is a result that hints at practical relevance in spite of the p-values~\cite{Ellis:2010:EGE}. Further investigation is required for these two atom candidates.

\begin{table}[th!]
\caption{Hypotheses Testing (correctness). Asterisks ($^{*}$) indicate a statistically significant difference. Using Bonferroni correction, a p-value is considered statistically significant if it is lower than 0.0021.}
 \centering
 {\scriptsize 
 \begin{tabular}{lrr}
   \toprule
Atom candidate & McNemar test &      Odds Ratio \\
\midrule
Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
Type Conversion           &   \textbf{$<$ 0.0001*} & 12.0 \\
Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.0 \\
Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
Pre-Increment              &  \textbf{0.0003*} &    16.0 \\
Omitted Curly Braces      &  \textbf{0.0009*} &  6.3 \\
Array Destructuring       &  0.0023 &     7.5 \\
Post-Increment              &  0.0225 &     5.5 \\
Array Spread              &  0.0490 &    3.3 \\
Object Spread             &  0.0636 &     2.8 \\
Property Access           &  0.0636 &     2.8 \\
Infix Operator Precedence &  0.1796 &     2.5 \\
Arrow Function            &  0.2891 &     3.0 \\
Arithmetic As Logic       &  0.3323 &  1.8 \\
Automated Semicolon Insertion  &  0.3915 &  1.4 \\
Logic As Control Flow     &  0.4421 &  1.5 \\
%Constant Variables        &  0.5000 &     0.0 \\
Implicit Predicate        &  0.6900 &  1.3 \\
Lack of Indentation With Braces   &  0.7266 &  1.7 \\
Conditional Operator      &  0.7905 &  1.3 \\
Dead Unreachable Repeated &  1.0000 &     1.5 \\
Repurposed Variables      &   1.0000 &     1.0 \\
\bottomrule
\end{tabular}
}
\label{tab:repeated:misunderstanding}
\end{table}

%-----------------------------------------------------------------------
\subsection{Time Analysis}\label{sec:repeated:time}

\subsubsection*{Exploratory Data Analysis}

In this section we investigate whether there is a trade-off between correctness and time required to predict program outputs, similarly to the analysis performed by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For this analysis, we consider only whether participants correctly predict the outputs of programs or not, independently of the presence or absence of atoms, and the time required to do so. The null hypothesis we investigate is that there is no significant difference in time between correct and incorrect predictions. Intuitively, we would expect correct responses to take longer, on the basis that participants would spend more time carefully analyzing the programs before providing an answer.

Table~\ref{tab:repeated:time} presents the mean and median times required by the participants to provide answers. The data is based on the 1648 correct and 1572 incorrect answers collected in the study. Contrary to our intuition, the mean time required by participants to provide an incorrect answer is 25\% greater than the time required to provide a correct one and the median time is 20\% greater. The standard deviations suggest that there is great variation in the times required by different participants, for different code snippets. Notwithstanding, the standard deviation for incorrect answers is 62.58\% greater when compared to the same statistic for correct answers.

\begin{table}[tb]
  \caption{Time in seconds required by the participants to correctly and incorrectly predict the outcomes of the programs.}\label{tab:repeated:time}
  \centering
  {\scriptsize 
  \begin{tabular}{lrrr}
    \toprule  
             & Mean  & Median & Std. Dev. \\
    \midrule
  Correct   & 35.44 & 24.36  & 38.43     \\
  Incorrect & 44.50 & 29.29  & 62.48    \\
  \bottomrule
  \end{tabular}
}
  \end{table}

\subsubsection*{Statistical Analysis}

We employ the non-parametric Mann-Whitney U test, also known as Wilcoxon's rank-sum test, to verify the aforementioned null hypothesis, i.e., whether the times for correct and incorrect answers can be considered identical. That test produces a p-value of \textbf{0.00000046}, which shows that the difference is statistically significant and we can reject the null hypothesis. To gauge the magnitude of that difference we employ Cliff's Delta as a measure of effect size. This produces a result of \textbf{0.1027}, which suggests that, although participants tend to provide correct predictions more quickly, they do that only slightly so.

%-----------------------------------------------------------------------
\section{Results of the Latin-Square Study}\label{sec:latin}

In the latin-square study, we estimate the impact of atoms considering the same two perspectives of the previous experiment: \emph{correctness} (number of wrong answers)
and \emph{time} (how long to provide a correct answer). In particular, in the time analysis, differently from the repeated  measures study, we compare the time required by participants to submit a correct response when analyzing code snippets with and without atom candidates. This study is based on the responses of 140 participants (a total of 70 replicas). All participants had taken at least some university course or hold a bachelor degree or equivalent. In addition, 21 participants hold a master's degree and three a doctorate degree. Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.  
Accordingly, we characterize the effect of atoms of confusion in JavaScript code taking into account the perceptions of both novice and experienced developers.

\subsection{Correctness Analysis}
% \castor{We need to be uniform about terminology. Are we going with ``misunderstanding rate analysis''? The problem is that it is not clear to me to which rate this refers.}

\subsubsection*{Exploratory Data Analysis}

As mentioned in Section~\ref{sec:meth:survey}, each participant in this experiment evaluated ten code snippets, from which five were in their obfuscated versions, whilst the other five contained clean versions of the code snippets (i.e., without the atom candidates). As in the repeated measures study, the participants should provide the expected outcomes of the code snippets. Differently from that study, in this one each participant only predicts the outcome of one version of each code snippet, either clean or obfuscated. As discussed, we collect information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}. 
Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the correctness experiment.

Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the obfuscated versions. In particular, the presence of the \emph{Comma Operator} atom exhibits the highest impact on misunderstanding. 
Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes. %introduce high degrees of confusion.
The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets. Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the clean versions of the code snippets are easier to evaluate correctly.

% \begin{table}[htbp]
% \caption{Difference in answer correctness between confusing and non-confusing pairs}
% \begin{center}
% \begin{small}
% \begin{tabularx}
% {{\linewidth}}{l p{1.5cm} p{1.1cm} p{1.1cm} p{1.2cm} }
% \textbf{Atom} & \textbf{\%Correct} & \textbf{\%Correct} \\
% &  \multicolumn{1}{l}{With AOC} \multicolumn{2}{l}{Without AOC}  & \Delta (\%)    \\
%  \hline
% Comma Operator & 40 & 93 & +132\%                 \\     
% Automatic Semicolon  Insertion & 46 & 97 & +110\% \\
% Post Increment & 69 & 91 & +31\%         \\
% Omitted Curly Braces & 67 & 83 & +23\%   \\ 
% Assignment as Value & 80 & 97 & +21\%    \\
% Implicit Predicate & 83 & 97 & +16\%     \\
% Logic as Control Flow & 59 & 68 & +15\%  \\
% Ternary Operator & 86 & 94 & +9\%        \\
% Pre-Increment & 71 & 76 & +7\%           \\
% Arithmetic as Logic & 91 & 90 & -1\%    \\
% \end{tabularx}
% \end{small}
% \end{center}
% \label{results_correctness}
% \end{table}

\begin{table}[htbp]
\caption{Summary of the correctness analysis}
\label{tab:difference-correctness}
\centering{
  {\scriptsize
  \begin{scriptsize}
\begin{tabular}{lrrr} \toprule
  Atom & Obfuscated & Clean & $\Delta$(\%)  \\ \midrule
  Comma Operator                &  28 &  65 & +132   \\ 
  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
  Post-Increment                &  48 &  64 & +33    \\ 
  Omitted Curly Braces          &  47 &  58 &  +23   \\
  Assignment as Value           &  56 &  68 &  +21   \\ 
  Implicit Predicate            &  58 &  68 &  +17   \\ 
  Logic as Control Flow         &  41 &  48 &  +17   \\ 
  Conditional Operator              &  60 &  66 &  +10   \\ 
  Pre-Increment                 &  50 &  53 &  +6    \\ 
  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule

\end{tabular}
\end{scriptsize}
}}
\end{table}

 %% Comma Operator          & 40 & 93  & +132 \\
 %% Post Increment          & 69 & 91  & + 31  \\
 %% Omitted Curly Braces    & 67 & 83  & +23 \\
 %% Assignment as Value     & 80 & 97  & +21 \\
 %% Implicit Predicate      & 83 & 97  & +16 \\
 %% Logic as Control Flow   & 59 & 68  & +15 \\
 %% Ternary Operator        & 86 & 94  & +9  \\
 %% Pre-Increment           & 71 & 76  & +7  \\
 %% Arithmetic as Logic     & 91 & 90  & +1  \\ \bottomrule

\begin{figure}[b!]
\noindent
 \centering
 \includegraphics[width=0.6\columnwidth]{images/wrong-answers-plot-1.pdf}
 \caption{Number of wrong answers of each subject.}
 \label{fig:boxplotcorrectness}
 \end{figure}% \castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou. Até mudei a figura. Não sei o que está havendo.}

% \rb{verificar a consistencia nas questoes de pesquisa mais especificas}

\subsubsection*{Statistical analysis}

We first use the \emph{Pearson's Chi-squared test}
to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (obfuscated and clean code). The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}. For five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate), the results indicate that the obfuscated versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.

We measure the effect size of the clean version of the code snippets into the answers' correctness using the \emph{Odds Ratio} (OR). The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}. 
In the table, we also report the confidence interval (CI) for the OR in the ``CI'' column. Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1. This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions. 
%Considering only the atoms for which there was a significant difference in correctness for confusing and clean versions and u
%Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), 
For four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large. For Implicit Predicate, it is medium. 
%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
For instance, when comparing clean and obfuscated code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}. This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding obfuscated version of the code snippet. Furthermore, with a 95\% confidence level, the odds ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version. If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio). This is the case for the four atom candidates in the lower part of the table. 
%In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a participant was actually more likely to misunderstand the clean version.

Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impacts correctness. The findings suggest that \emph{participant experience} impacts the results related to correctness (p-value $<$ 0.001 and $\chi^2 = 24.18$).
Figure~\ref{fig:correctness-over-experience} summarizes the number of correct and incorrect answers over the four experience groups. Even though the presence of atoms of confusion reduces the number of correct answers for all \emph{experience groups}, this effect is not uniform. For instance, novice developers (under one year of experience) provide a higher number of wrong answers for the obfuscated version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the obfuscated version of the code snippets. It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code. We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience (p-value = 0.007 and $\chi^2$ = 7.08). Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers (p-value = 0.236 and $\chi^2$ = 5.54).

% \castor{I'd like to have the numbers for the results above. What are the p-values, regression coefficients, etc.?}

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{images/correctness-by-experience}
  \caption{Impact of atoms of confusion on the correctness (over the four participant experience groups)}
  \label{fig:correctness-over-experience}
\end{figure}

%% \begin{mh}
%%   The results of our exploratory data analysis and   hypothesis testing suggest that the atom candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, Assignment as Value, and Implicit Predicate %, and Ternary Operator
%%   introduce some degree of misunderstanding in JavaScript code. %Based on the results of this study, 
%%   Therefore, they can be deemed atoms of confusion for JavaScript.
%% \end{mh}

%% \begin{mh}
%% {\color{red}Comparing the results of different studies,
%%   the impact of atoms candidates on source code
%%   (mis)understanding differ according to the programming
%%   language.}
%% \end{mh}
%% Regarding the first question we 
%% address in the survey (\emph{Do code snippets that contain atoms of confusion produce a higher error rate than snippets where the atom is removed?}), we found evidence that the atoms of confusion lead programmers to misunderstand JavaScript code. We also realized that just one atom whose correction has a non-significant improvement in the percentage of correct answers---we found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of ten atoms we consider in the survey).

\subsection{Time Analysis}

\subsubsection*{Exploratory Data Analysis}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions. We do not consider cases where the participants made mistakes. Seven atom candidates required more time from participants to predict a correct response. For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, require less time for the participants to predict the answer. In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was   
29.06\% and 38.19\% than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.

\begin{table}[htbp]
\caption{Time in seconds to submit a correct answer}
\centering{
  {\scriptsize
  \begin{scriptsize}
\begin{tabular}{lrrr} \toprule
Atom & Obfuscated  & Clean & $\Delta$(\%) \\ \midrule
Comma Operator & 87.67 & 20.84 & -76.23 \\ 
Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
Post-Increment & 28.70 & 25.67 & -10.56 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
Conditional Operator & 41.80 & 39.34 & -5.90 \\ 
Pre-Increment & 30.71 & 42.45 & 38.19 \\  
Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
\end{tabular}
\label{tab:difference-time-taken}
\end{scriptsize}
}}
\end{table}

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
 \centering
 {\scriptsize
 \begin{tabular}{lrrr|rr}
   \toprule
Atom  & Chi-square & Odds          & Confidence & Mann-Whitey& Cliff's\\ 
 &  test &Ratio          & Interval &  U test & Delta \\  \midrule
Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
Automatic Semicolon Insertion & \textbf{$<$ 0.0001*} & 39.33 & (9.21, 356.62) & 0.0980 & -0.16 \\ 
Post-Increment & \textbf{0.0015*} & 4.83 & (1.73, 15.74) & 0.1201 & 0.15 \\ 
Omitted Curly Braces & 0.0510 & 2.35 & (1.00, 5.77) & 0.9253 & -0.01 \\ 
Assignment as Value & \textbf{0.0035*} & 8.39 & (1.81, 79.12) & 0.8268 & 0.02 \\ 
Implicit Predicate & \textbf{0.0112*} & 6.95 & (1.46, 66.56) & \textbf{0.0029*} & -0.29 \\ 
Logic as Control Flow & 0.2920 & 1.54 & (0.73, 3.28) & \textbf{$<$ 0.0001*} & -0.39 \\ 
Conditional Operator & 0.1590 & 2.73 & (0.74, 12.57) & 0.2407 & -0.12 \\ 
Pre-Increment & 0.7015 & 1.25 & (0.55, 2.85) & \textbf{0.0062*} & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & \textbf{0.0172*} & 0.23 \\ 
    \bottomrule
 \end{tabular}}
 \label{tab:hypothesis-testing}
 \end{table*}

\subsubsection*{Statistical analysis}
We use the \emph{Mann-Whitney U test} to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of clean and obfuscated versions of a code snippet.
%, regardless of evaluating the confusing or clean version of the code.
The results of Table~\ref{tab:hypothesis-testing}
suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the obfuscated code version. For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the obfuscated version. We also computed the effect size using Cliff's Delta statistic. In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.  Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.

%% \begin{mh}
%%   The results suggest that developers tend to spend
%%   more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate. Conversely, the code snippets with the atom candidates Arithmetic as Logic and Pre-Increment demand less time to predict the correct answers. 
%% \end{mh}

%%  Comma Operator        & 60 & 21  & -65  \\
%% Logic as Control Flow & 85 & 49  & -42  \\
%% Implicit Predicate    & 33 & 24  & -27  \\
%% Omitted Curly Braces  & 43 & 31  & -27  \\
%% Assignment as Value   & 53 & 49  & -7   \\
%% Ternary Operator      & 42 & 42  & 0    \\
%% Post Increment        & 27 & 27  & 0    \\
%% Arithmetic as Logic   & 29 & 36  & +24  \\
%% Pre-Increment         & 34 & 49  & +44  \\

% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}

\section{Results of the Interview Study}
\label{sec:interview-results}

@@

%% \rb{acho que essa separa\c c\~{a}o em dois rounds est\'{a}
%%   gerando uma complexidade desnecess\'{a}ria. vou tentar
%%   remover isso}.

%In this section, we present the results of the 
% interviews with the 15 practitioners who agreed to participate in the study. 
A total of 15 practitioners took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the both experiments. We first presented them pairs of clean and obfuscated code snippets and then asked them to discuss their preferences towards one version. We did not indicate in any way whether any version was assumed to be confusing or not. 
% ask the participants to discuss their preferences (i.e., code with or without atom).

%% In the second round of interviews, ten of the fifteen developers
%% who participated in the first round were shown another 8 atom candidates to atom of confusion whose frequency we observed during the repository mining phase. The same procedure of showing code that behaves exactly the same and surveying for how easy to understand each snippet was applied in this phase.

@@

\subsubsection*{The Participants' perceptions of the atom candidates}

@@

\subsubsection*{Confusion in JavaScript Code}

@@

% citar uninitialized objects' types. floating point precision issues, 'this', type coercion

% \subsection{Common Misunderstandings About the Language}

% TODO: I don't see that our results support these claims.
% Let us comment for now.

%% Six practitioners regarded the JavaScript language as easy to learn, while four
%% participants mentioned that the abundance of frameworks,
%% coupled with the abstractions they provide, make it so that one
%% can develop quickly in the language. However, as mentioned by one
%% of the interviewees, ``\emph{Developers who have started
%% programming in JavaScript might miss out on important lower-level
%% aspects, such as garbage collection, memory management,
%% and process management.}''.

>>>>

\section{Interview Study}\label{sec:method:interview}

To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2). We also ask each participant if they know of any other JavaScript-specific construct or idiom they think is likely to make the code hard to understand. % (research question RQ3). 
In this section we detail the protocol we followed to conduct the interviews and to analyze the results.

% Nós realizamos entrevistas semi-estruturas com o objetivo de identificar a percepção dos desenvolvedores com experiência em JavaScript sobre algumas questões relacionadas à compreensão de código em JavaScript. Assim, nesta Seção nós descrevemos os procedimentos adotados para selecionar os participantes para as entrevistas e detalhamos como as entrevistas foram conduzidas. Além disso, detalhamos como os resultados das entrevistas foram analisados.

\subsection{Participant Selection} We invite the participants of the interviews using a snowballing technique. That is, starting from our network of contacts, we invite an initial set of candidates to take part in our experiments. From this initial list, we ask for an indication of additional candidates. Our main selection criterion is that all participants should have been working with JavaScript in their daily professional activities. We invited a total of 17 developers, and 15 of them agreed to participate.   
%We conducted the interviews for a period of two weeks.

\subsection{Interview Process} We conduct semi-structured interviews using web conferencing software. We record all the interviews with the consent of the participants. On average, the interviews last 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes. Two of the researchers conduct the interviews, and a third one listens to all the recordings to cross-validate the collected data. The interviews have three main parts. In the first one, we ask the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years, and other programming languages they have worked with. Table \ref{pinterview} summarizes this demographic information.

\begin{table}[htb!]

  \centering
  \caption{Demographic information of the participants}
\begin{scriptsize}  
\begin{tabular}{clcl}
\toprule
ID & Education & JS Experience & Other Languages \\ \midrule 
P1 & BSc Degree & 9 years & Java, PHP, C, Go  
\\ 
P2 & HS Degree & 3 years & Python, Go, Dart, Lua, C++, C\#
\\ 
P3 & BSc Degree & 4 years & Java, C
\\ 
P4 & Undergraduate & 3 years & Python, C, C++, Java, Go
\\ 
P5 & Master Student & 3 years & Python, SQL
\\ 
P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\#
\\ 
P7 & BSc Degree  & 6 years & C, C++, Java, Assembly, Kotlin
\\ 
P8 & PhD Degree & 2 years & Java, Python
\\ 
P9 & BSc Degree  & 5 years & PHP
\\ 
P10 & Master Student & 4 years & Java, C\# and Python
\\ 
P11 & Master Student & 4 years & Java, Erlang, C\#, Cobol
\\ 
P12 & Master Degree & 1 year & C
\\ 
P13 & Master Degree & 13 years & Java, PHP
\\ 
P14 & Master Student & 3 years & C, Python, Ruby
\\ 
P15 &  BSc degree  & 2 years & Java, PHP
\\ \bottomrule
\end{tabular}
\end{scriptsize}
    \label{pinterview}
\end{table}

In the second part of the interview our aim is to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regard as innately confusing. This allows us to identify potential atom candidates that are more specific to the JavaScript language. Examples of questions we explore in this section include: \emph{Does JavaScript favor developers to produce code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as especially confusing?}

In the third part of the interview, participants are shown pairs of programs that were used in the experiments, where each pair contains a clean and an obfuscated version of the same atom candidate. For this part we only use atoms that appear in both experiments. The participants 
are asked to evaluate which version of the code is easier to understand. To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation. Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.

\subsection{Interview Analysis}

We first transcribe each interview and then examine the broad distribution of the answers. Our goal is to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular. We follow-up with an open-coding procedure, highlighting the main themes and quoting the answers of the participants. We present these results in Section~\ref{sec:interview-results}.

\subsection{Results}\label{sec:interview-results}

@@

A total of 15 practitioners took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in the both experiments. We first presented them pairs of clean and obfuscated code snippets and then asked them to discuss their preferences towards one version. We did not indicate in any way whether any version was assumed to be confusing or not.

@@

\subsubsection{The Participants' perceptions of the atom candidates}

@@

\subsubsection{Confusion in JavaScript Code}

@@
